---
title: "3.Environmental"
author:
  - Asa12138
date: Mon Apr  8 21:56:49 2024
link-citations: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
path <- "/Users/asa/Documents/R/CRISPR_analysis/analysis"
knitr::opts_chunk$set(eval=TRUE, #run code in chunks (default = TRUE)
                       echo = TRUE, #whether to include the source code in the output
                       error = TRUE, #Whether to include error information in the output
                       warning = FALSE, #Whether to include warnings in the output (default = TRUE)
                       message = TRUE, #whether to include reference information in the output
                       cache=TRUE, #whether to cache
                       collapse = FALSE # output in one piece
                       )
knitr::opts_knit$set(root.dir = path)
```

Install and import all dependent packages, data import:
```{r import,echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
source("./R_config.R")
```

# Pre_CCF_res

## microbiome
1. oral：/data/home/jianglab/share/oral_fna/spacer/ 1225个genome
2. gut3: /data/home/jianglab/share/gut2/gut3/output 3909个genome
3. skin: 
    /data/home/jianglab/share/skin_microbiome/MAG_output/output   
    /data/home/jianglab/share/skin_microbiome/ISO_skin/output
4. glacier:
    /data/home/jianglab/share/glacier/iso_output/output
    /data/home/jianglab/share/glacier/MAG_output/output
5. soil:  /share/home/jianglab/shared/metagenome/genomic_earth_micro/soil/output
6. marine: /share/home/jianglab/shared/metagenome/genomic_earth_micro/marine/output

```{r eval=FALSE}
library(iCRISPR)
library(pcutils)
library(dplyr)
input_dirs=c(
  oral_fna="/data/home/jianglab/share/oral_fna/spacer/",
  gut3="/data/home/jianglab/share/gut2/gut3/output",
  MAG_output="/data/home/jianglab/share/skin_microbiome/MAG_output/output",
  ISO_skin="/data/home/jianglab/share/skin_microbiome/ISO_skin/output",
  iso_output="/data/home/jianglab/share/glacier/iso_output/output",
  MAG_glacier="/data/home/jianglab/share/glacier/MAG_output/output",
  soil="/share/home/jianglab/shared/metagenome/genomic_earth_micro/soil/output",
  marine="/share/home/jianglab/shared/metagenome/genomic_earth_micro/marine/output"
)

for (i in 1:8) {
  work_dir=file.path("/share/home/jianglab/shared/ouxinyi2liuzhen/bacteria24/env_crispr",names(input_dirs)[i])
  if(!dir.exists(work_dir))dir.create(work_dir)
  setwd(work_dir)
  res=multi_pre_CCF_res(input_folder = input_dirs[i],output_folder = NULL,threads=1)
  save(res,file = "all_crispr.rda")
  all_fa_res=get_spacer_fa(res,evidence_level = 4)
  save(all_fa_res,file ="all_level4_spacer.rda")
  pcutils::write_fasta(all_fa_res,"all_level4_spacer.fa")
}
```

/data/home/jianglab/share/glacier/MAG_output/output下报错，试试

/data/home/jianglab/share/glacier/MAG_output/output/MAG_1611DDS_maxbin2_bin_0
/data/home/jianglab/share/glacier/MAG_output/output/MAG_1611DDS_metabat2_bin_6_002
```{bash}
scp -P 45777 -r pengchen@10.73.29.10:/data/home/jianglab/share/glacier/MAG_output/output/MAG_1611DDS_maxbin2_bin_0 ~/Documents/R/CRISPR_analysis/data/
scp -P 45777 -r pengchen@10.73.29.10:/data/home/jianglab/share/glacier/MAG_output/output/MAG_1611DDS_metabat2_bin_6_002 ~/Documents/R/CRISPR_analysis/data/
```

```{r}
pre_CCF_res(input_folder = "../data/MAG_1611DDS_metabat2_bin_6_002",output_folder = NULL)
```

```{r}
input_dirs=c(
  oral_fna="/data/home/jianglab/share/oral_fna/spacer/",
  gut3="/data/home/jianglab/share/gut2/gut3/output",
  MAG_output="/data/home/jianglab/share/skin_microbiome/MAG_output/output",
  ISO_skin="/data/home/jianglab/share/skin_microbiome/ISO_skin/output",
  iso_output="/data/home/jianglab/share/glacier/iso_output/output",
  MAG_glacier="/data/home/jianglab/share/glacier/MAG_output/output",
  soil="/share/home/jianglab/shared/metagenome/genomic_earth_micro/soil/output",
  marine="/share/home/jianglab/shared/metagenome/genomic_earth_micro/marine/output"
)
for (i in 1:8) {
  work_dir=file.path("/share/home/jianglab/shared/ouxinyi2liuzhen/bacteria24/env_crispr",names(input_dirs)[i])
  file.rename(file.path(work_dir,paste0(names(input_dirs)[i],"all_crispr.rda")),file.path(work_dir,paste0(names(input_dirs)[i],"_all_crispr.rda")))
}
```

```{r}

for (i in c("ISO_skin/ISO_skin_all_crispr.rda",
"gut3/gut3_all_crispr.rda ",
"oral_fna/oral_fna_all_crispr.rda", 
"MAG_glacier/MAG_glacier_all_crispr.rda",
"iso_output/iso_output_all_crispr.rda",
"soil/soil_all_crispr.rda", 
"MAG_output/MAG_output_all_crispr.rda",
"marine/marine_all_crispr.rda")) {
  print(i)
  file1=paste0("/share/home/jianglab/shared/ouxinyi2liuzhen/bacteria24/env_crispr/",i)
  system(paste0("scp -P 45777 -r pengchen@10.73.29.10:",file1," ~/Documents/R/CRISPR_analysis/data/env/"))
}

```

```{r}
load_all("~/Documents/R/CRISPR/iCRISPR/")
list.files("../data/env/",pattern = "all_crispr.rda",full.names = T)->all_crispr_files
```

```{r}
tmp1=tmp2=tmp3=list()
for (i in names(env_match)) {
  print(i)
  load(paste0("../data/env/",i,"_all_crispr.rda"))
  tmp1[[env_match[i]]]=data.frame(env=env_match[i],get_crispr_info(res),row.names = NULL)
  tmp2[[env_match[i]]]=lapply(res, \(i)i$CRISPR)%>%do.call(rbind,.)%>%
    data.frame(env=env_match[i],.,row.names = NULL)
  tmp3[[env_match[i]]]=lapply(res, \(i)i$Cas)%>%do.call(rbind,.)%>%
    data.frame(env=env_match[i],.,row.names = NULL)
  #Cas只需要管Cas id对应的type
  distinct_at(tmp3[[env_match[i]]],c(1:4,6,7))->tmp3[[env_match[i]]]
}
do.call(rbind,tmp1)%>%write.csv(file = "3.Environmental/env_crispr_info.csv",row.names = F)
do.call(rbind,tmp2)%>%write.csv(file = "3.Environmental/env_array_info.csv",row.names = F)
do.call(rbind,tmp3)%>%write.csv(file = "3.Environmental/env_cas_info.csv",row.names = F)
```


```{r}
env_match=c(
  oral_fna="oral",
  gut3="gut",
  MAG_output="skin_mag",
  ISO_skin="skin_iso",
  iso_output="glacier_iso",
  MAG_glacier="glacier_mag",
  soil="soil",
  marine="marine"
)

cas_type_list=list()
for (i in names(env_match)) {
  print(i)
  load(paste0("../data/env/",i,"_all_crispr.rda"))
  cas_type_list[[env_match[i]]]=summary_cas_type(res,each_genome = F)
}
rm(res)

cas_type_list[["skin"]]<-combine(cas_type_list[["skin_mag"]],cas_type_list[["skin_iso"]])
cas_type_list[["glacier"]]<-combine(cas_type_list[["glacier_mag"]],cas_type_list[["glacier_iso"]])
        
save(cas_type_list,file = "3.Environmental/cas_type_list.rda")
```

```{r}
load("3.Environmental/cas_type_list.rda")

all_type=do.call(rbind,cas_type_list)
all_type=c(unique(all_type[,1]),unique(all_type[,2]))
all_type_col=get_cols(all_type)

plist=list()
for (i in names(cas_type_list)) {
  print(i)
  plot.cas_type(cas_type_list[[i]])+
    scale_fill_manual(values = all_type_col)+
    ggtitle(i)+xlim(0,3.6)->plist[[i]]
}

plot_grid(plotlist = plist[1:8],nrow = 2)
ggsave("3.Environmental/env_cas_type.pdf",width = 18,height = 8)

envs=c("oral","gut","skin","glacier","soil","marine")
plot_grid(plotlist = plist[envs],nrow = 2)
ggsave("3.Environmental/env_cas_type2.pdf",width = 13,height = 8)
```


## Tree

```{bash}
scp -P 45777 -r pengchen@10.73.29.10:/share/home/jianglab/shared/ouxinyi2liuzhen/bacteria24/env_crispr/env_tree_source ~/Documents/R/CRISPR_analysis/analysis/3.Environmental/
```

```{r}

#修改一些颜色
phylum_color=setNames(c("#8DD3C7","#FAAB78","#C37B89","#FB8072","#6E85B7","grey","#FCCDE5","#5D534A","#B2A4FF","#986D8E","#E78AC3","#A6D854","#FFD92F","#E5C494","#416D19","#B5CDA3"),c("Bacillota","Pseudomonadota","Thermoproteota","Chloroflexota","Bacteroidota","Others","Actinomycetota","Verrucomicrobiota","Campylobacterota","Cyanobacteriota","Spirochaetota","Euryarchaeota","Planctomycetota","Thermodesulfobacteriota","Thermotogota","Deinococcota"))


library(pctax)
library(ggtree)
library(ggtreeExtra)
read_delim("3.Environmental/env_tree_source",delim  = "\t")->env_tree_source
f_tax=strsplit2(env_tree_source$spacer_lineage,split = ";",
                                         colnames = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))
f_tax2=f_tax
f_tax2=pctax::pre_tax_table(f_tax2)
f_tax2=before_tree(f_tax2)
env_tree_source$species=f_tax2$Species

spe_spacer_count=count(env_tree_source,environment,species,name = "count")
spe_spacer_count=spe_spacer_count%>%mutate(log_count=log(count),log2_count=log2(count))

ann_tree(f_tax2)->tree

p=easy_tree(tree,highlight = "Phylum",add_tiplab = F,topN = 9,pal = phylum_color)

p1=p+geom_fruit(
    geom = geom_tile,
    data = spe_spacer_count,
    mapping = aes(
        y=species,
        x=environment,
        fill=environment,
        alpha=log_count
    ),
    axis.params = list(axis ="x",text.size=3,text.angle=0,hjust=1)
)+scale_fill_manual(values = c('oral'="#668B8B",'desert'="mistyrose4",'glacier'="#9370DB",'vagina'="dodgerblue3",
                               'skin'="mistyrose3",'archaea'="#E82E6F",'gut'="#FF904F",'marine'="#9CFAFF",'soil'="gold"))+
    scale_alpha(range = c(0.2,1))

open_tree(p1,10)%>%rotate_tree(-90)

ggsave("3.Environmental/env_tree_source.pdf",width = 10,height = 8)
```


Tree2, 画一个物种有无的树，因为上面的只展示有无spacer，没有spacer的可能是没有该物种，也可能有物种没spacer

```{bash}
scp -P 45777 -r pengchen@10.73.29.10:/share/home/jianglab/shared/ouxinyi2liuzhen/bacteria24/env_crispr/env_total_taxon_tree 3.Environmental/
```


```{r}
load_all("../../pctax/pctax/")
read_delim("3.Environmental/env_total_taxon_tree",delim  = "\t")->env_total_taxon_tree
f_tax=strsplit2(env_total_taxon_tree$lineage,split = ";",
                                         colnames = c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))
f_tax2=f_tax
f_tax2=pctax::pre_tax_table(f_tax2)
f_tax2=before_tree(f_tax2)
env_total_taxon_tree$species=f_tax2$Species

spe_count=count(env_total_taxon_tree,environment,species,name = "count")

ann_tree(f_tax2)->tree_all

p=easy_tree(tree_all,highlight = "Phylum",add_tiplab = F,topN = 9,pal = phylum_color)

p1=p+geom_fruit(
    geom = geom_tile,
    data = spe_count,
    mapping = aes(
        y=species,
        x=environment,
        fill=environment
    ),
    axis.params = list(axis ="x",text.size=3,text.angle=0,hjust=1)
)+scale_fill_manual(values = c('oral'="#668B8B",'desert'="mistyrose4",'glacier'="#9370DB",'vagina'="dodgerblue3",
                               'skin'="mistyrose3",'archaea'="#E82E6F",'gut'="#FF904F",'marine'="#9CFAFF",'soil'="gold"))

open_tree(p1,10)%>%rotate_tree(-90)

ggsave("3.Environmental/env_total_taxon_tree.pdf",width = 10,height = 8)


#绘制之前的spacer树，但是注释改为物种是否存在看看
p=easy_tree(tree,highlight = "Phylum",add_tiplab = F,topN = 9,pal = phylum_color)

p1=p+geom_fruit(
    geom = geom_tile,
    data = spe_count,
    mapping = aes(
        y=species,
        x=environment,
        fill=environment
    ),
    axis.params = list(axis ="x",text.size=3,text.angle=0,hjust=1)
)+scale_fill_manual(values = c('oral'="#668B8B",'desert'="mistyrose4",'glacier'="#9370DB",'vagina'="dodgerblue3",
                               'skin'="mistyrose3",'archaea'="#E82E6F",'gut'="#FF904F",'marine'="#9CFAFF",'soil'="gold"))

open_tree(p1,10)%>%rotate_tree(-90)
ggsave("3.Environmental/env_tree_source_taxon.pdf",width = 10,height = 8)

```

